---
title: "Datasets and shipped artifacts"
author: "elcf4R package"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Datasets and shipped artifacts}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(elcf4R)

data("elcf4r_iflex_example", package = "elcf4R")
data("elcf4r_iflex_benchmark_results", package = "elcf4R")
data("elcf4r_storenet_example", package = "elcf4R")
data("elcf4r_storenet_benchmark_results", package = "elcf4R")
data("elcf4r_lcl_example", package = "elcf4R")
data("elcf4r_lcl_benchmark_results", package = "elcf4R")
data("elcf4r_refit_example", package = "elcf4R")
data("elcf4r_refit_benchmark_results", package = "elcf4R")
```

# Overview

`elcf4R` now supports four household-oriented public data sources through a
common normalized panel schema:

- `elcf4r_read_iflex()`
- `elcf4r_read_storenet()`
- `elcf4r_read_lcl()`
- `elcf4r_read_refit()`

The package also ships compact example panels and saved benchmark results so
the main vignettes run without external downloads. Raw source files stay in
`data-raw/` and are not redistributed through the package unless a compact
derived artifact has been explicitly built and saved.

Two additional unshipped scaffolds are also available:

- `elcf4r_download_ideal()` / `elcf4r_read_ideal()` for aggregate-electricity
  hourly summaries from IDEAL.
- `elcf4r_download_gx()` / `elcf4r_read_gx()` for the GX
  transformer/community-level dataset.

# Supported dataset matrix

The current dataset surface is:

| Dataset | Reader | Resolution | Temperature in normalized panel | Shipped example | Shipped benchmark |
|:--|:--|:--|:--|:--|:--|
| iFlex | `elcf4r_read_iflex()` | hourly | yes | `elcf4r_iflex_example` | `elcf4r_iflex_benchmark_results` |
| StoreNet (`H6_W`) | `elcf4r_read_storenet()` | 1 minute | optional, source-dependent | `elcf4r_storenet_example` | `elcf4r_storenet_benchmark_results` |
| Low Carbon London | `elcf4r_read_lcl()` | 30 minutes | no | `elcf4r_lcl_example` | `elcf4r_lcl_benchmark_results` |
| REFIT | `elcf4r_read_refit()` | user-selected resample | no | `elcf4r_refit_example` | `elcf4r_refit_benchmark_results` |
| ELMAS | not part of the common household reader set | hourly | no | `elcf4r_elmas_toy` | none |

All four household readers return the same core columns:

- `dataset`
- `entity_id`
- `timestamp`
- `date`
- `time_index`
- `y`
- `temp`
- `dow`
- `month`
- `resolution_minutes`

Dataset-specific metadata columns are preserved when available.

# Scaffolded, unshipped datasets

`IDEAL` and `GX` are intentionally documented separately from the core shipped
household matrix.

| Dataset | Helper surface | Level | Current scaffold scope | Shipped example | Shipped benchmark | Licence note |
|:--|:--|:--|:--|:--|:--|:--|
| IDEAL | `elcf4r_download_ideal()`, `elcf4r_read_ideal()` | household | aggregate-electricity hourly summaries from `auxiliarydata.zip` | no | no | the current Edinburgh DataShare record states `CC BY 4.0` |
| GX | `elcf4r_download_gx()`, `elcf4r_read_gx()` | transformer/community | SQLite or flat-export normalization to the common panel schema | no | no | treat licence terms as dataset-record specific and recheck before redistribution |

Notes:

- IDEAL support in this release is limited to aggregate electricity and does
  not attempt to parse the raw 1 Hz stream.
- GX is not an individual-household dataset. It is useful as a secondary
  benchmark source for weather and community-level demand behavior, but it is
  not folded into the package's core household benchmark claims.

# Shipped example panels

The shipped examples are small normalized panels intended for package examples
and vignette code.

```{r}
example_sizes <- data.frame(
  dataset = c("iflex", "storenet", "lcl", "refit"),
  rows = c(
    nrow(elcf4r_iflex_example),
    nrow(elcf4r_storenet_example),
    nrow(elcf4r_lcl_example),
    nrow(elcf4r_refit_example)
  ),
  ids = c(
    length(unique(elcf4r_iflex_example$entity_id)),
    length(unique(elcf4r_storenet_example$entity_id)),
    length(unique(elcf4r_lcl_example$entity_id)),
    length(unique(elcf4r_refit_example$entity_id))
  )
)

example_sizes
```

These objects can be passed directly to:

- `elcf4r_build_daily_segments()`
- `elcf4r_build_benchmark_index()`
- `elcf4r_benchmark()`

# Shipped benchmark result datasets

Each supported household dataset now has a saved benchmark-result object built
from a fixed local cohort and a deterministic rolling-origin design.

```{r}
benchmark_summary <- do.call(
  rbind,
  list(
    transform(
      aggregate(cbind(nmae, nrmse, smape, mase) ~ method,
                data = elcf4r_iflex_benchmark_results,
                FUN = function(x) round(mean(x, na.rm = TRUE), 4)),
      dataset = "iflex"
    ),
    transform(
      aggregate(cbind(nmae, nrmse, smape, mase) ~ method,
                data = elcf4r_storenet_benchmark_results,
                FUN = function(x) round(mean(x, na.rm = TRUE), 4)),
      dataset = "storenet"
    ),
    transform(
      aggregate(cbind(nmae, nrmse, smape, mase) ~ method,
                data = elcf4r_lcl_benchmark_results,
                FUN = function(x) round(mean(x, na.rm = TRUE), 4)),
      dataset = "lcl"
    ),
    transform(
      aggregate(cbind(nmae, nrmse, smape, mase) ~ method,
                data = elcf4r_refit_benchmark_results,
                FUN = function(x) round(mean(x, na.rm = TRUE), 4)),
      dataset = "refit"
    )
  )
)

benchmark_summary[, c("dataset", "method", "nmae", "nrmse", "smape", "mase")]
```

These shipped benchmark tables are poster-style artifacts. They are not
intended to replace full local benchmarking on the raw datasets.

# Rebuilding the shipped artifacts

Each shipped dataset is reproducible from a `data-raw/` script:

- `data-raw/elcf4r_iflex_subsets.R`
- `data-raw/elcf4r_iflex_benchmark_results.R`
- `data-raw/elcf4r_storenet_artifacts.R`
- `data-raw/elcf4r_lcl_artifacts.R`
- `data-raw/elcf4r_refit_artifacts.R`

The general pattern is:

1. Place the original raw files in `data-raw/`.
2. Read them through `elcf4r_read_*()`.
3. Build a normalized day index with `elcf4r_build_benchmark_index()`.
4. Save a compact example panel.
5. Run `elcf4r_benchmark()` on a fixed cohort and save the result table.

This keeps the package lightweight while making the shipped examples and
benchmark summaries reproducible.

# Example: daily segments from a shipped panel

```{r}
iflex_segments <- elcf4r_build_daily_segments(
  elcf4r_iflex_example,
  carry_cols = c("dataset", "participation_phase", "price_signal")
)

dim(iflex_segments$segments)
head(iflex_segments$covariates[, c("entity_id", "date", "temp_mean", "price_signal")])
```

# Example: rerun a tiny benchmark locally

```{r}
tiny_index <- elcf4r_build_benchmark_index(
  elcf4r_lcl_example,
  carry_cols = "dataset"
)

tiny_benchmark <- elcf4r_benchmark(
  panel = elcf4r_lcl_example,
  benchmark_index = tiny_index,
  methods = c("gam", "kwf"),
  cohort_size = 1,
  train_days = 10,
  test_days = 2,
  include_predictions = FALSE
)

tiny_benchmark$results[, c("entity_id", "method", "test_date", "nmae", "mase")]
```
